Automatically generated NE tagged corpora for English and Hungarian

### 2012-07-02 paper/jeju_presentation.tex
Final version of the META-SHARE descriptor and penultimate version of the Jeju presentation.

\newcommand{\vitem}{\vfill \item}
\newcommand{\nyil}{$\rightarrow$\ }
\newcommand{\nagytilde}{$\sim$}

@@

The standard datasets are

\begin{itemize}
\vitem highly domain-specific
\vitem restricted in size
\end{itemize}
\vfill

@@

To get a bigger training corpus:

\begin{itemize}
\vitem merging datasets \nyil problems of combining different tagsets
\vitem manual annotation \nyil not cost-effective
\end{itemize}
\vfill

@@

\begin{frame}{Reducing the annotation cost}

@@

Reducing the annotation cost:

\begin{itemize}
\item to generate the resource automatically 
\item to use collaboratively constructed resources (e.g.~Wikipedia, DBpedia)
\end{itemize}

@@

\begin{nix}
combining these approaches: \\ automatically generating NE tagged corpora from Wikipedia
\end{nix}

@@

Improving the accuracy of NER with WP:

\begin{itemize}
\item extracting gazetteers containing person names, locations and organizations 
\item extracting the knowledge embodied in WP and incorporating it in NER learning as features
\item automatic creation of training data from WP
\begin{itemize}
\item by using the category structure of WP
\item by using the category head nouns and the opening sentences of articles
\end{itemize} 
\end{itemize}

@@

\begin{frame}{Our approach}

\begin{nix}
mapping the DBpedia ontology classes to standard NE tags, and assign them to WP entities
\end{nix}

@@

our method is largely language-independent \nyil sufficiently large corpora for less resourced languages can be built

@@

\begin{nix}
this is the first automatically annotated corpus for Hungarian 
\end{nix}

@@

\subsection*{Motivation}
\begin{frame}{Motivation}
  \begin{itemize}
  \vitem Our goal is to create a large NE annotated corpus, automatically
        generated from WP articles.
  \vitem Our method is similar to Nothman et al. (2008)
  \vitem The process consists of four steps:
    \begin{enumerate}
    \vitem Classify WP articles into entity classes.
    \vitem \textit{Parse WP and split articles into sentences.}
    \vitem Label named entities in the text.
    \vitem Select the sentences for inclusion in the corpus.
    \end{enumerate}
  \end{itemize}
\end{frame}

\subsection{Articles to Entities}
\begin{frame}{Articles to Entities}
  Our method is based on thinking of WP articles as NEs.
  \begin{itemize}
  \vitem Usual approach: semi-supervised learning (infoboxes, first sentence)
  \vitem Our approach: create entity list from DBpedia
    \begin{itemize}
    \vitem OWL ontology with 320 classes
    \vitem We take the most specific class
    \vitem Manual mapping to CoNLL categories that exploits inheritence
    \end{itemize}
  \vitem Decreased coverage, increased precision
  \end{itemize}
\end{frame}

\subsection{Named Entity Tagging}
\begin{frame}{Named Entity Tagging}
  The entity list from DBpedia is used for NE tagging.
  \begin{itemize}
  \vitem Entity identification:
    \begin{itemize}
    \vitem Look up linked pages in the entity list
    \vitem References to the current page, if in the list
    \end{itemize}
  \vitem NE tagging: from the DBpedia--CoNLL category mapping
  \vitem Exceptions:
    \begin{itemize}
    \vitem Not entities: common nouns, date expressions, etc.
    \vitem Unmarked entities: link to already identified entities
    \vitem Special cases: derived words, first word in a sentence, etc.
    \end{itemize}
  \end{itemize}
\end{frame}

\subsection{Sentence Filtering}
\begin{frame}{Sentence Filtering}
  We only include fully tagged sentences in the corpus.
  \begin{itemize}
  \vitem Sentences that contain unknown links / uppercase words
  \vitem To filter incomplete sentences:
    \begin{itemize}
    \vitem Sentences without an EOS punctuation mark
    \vitem Sentences without a predicate (verb)
    \end{itemize}
  \end{itemize}
  \vfill
\end{frame}

@@

%TODO: the "Hungarian Corpus" phrase must appear in the frame title 
\begin{frame}{Differences from English}
  \begin{itemize}
  \vitem Language processing tools
    \begin{itemize}
    \vitem Inflectional morphology: different set of tools were needed 
    \vitem We developed a flexible framework that enables the plugging of any NLP tool
    % TODO: what kind of linguistic tools are needed: table?
    \end{itemize}
  \vitem Wikipedia size and coverage
    \begin{itemize}
    \vitem Hungarian is medium-size language: \nagytilde 200k articles vs.~\nagytilde 4M English pages
    \vitem Cross-references may point to English pages
    \vitem Templates are in Hungarian: per-language template configuration
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Differences from English -- cont.}
  \begin{itemize}
  \vitem DBpedia coverage
    \begin{itemize}
    \vitem Internationalized chapters for big languages only
    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
      \begin{itemize}
      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
      \vitem Hungarian entities without an English page are missing:
             the NER language model might not be accurate
      \vitem Future work: extend the list with the missing Hungarian entities
      \end{itemize}
    \end{itemize}
  \vitem NE conventions
    \begin{itemize}
    \vitem Szeged NER guideline: similar to CoNLL
    \vitem Common nouns in Hungarian: names of languages, religions, months, etc. \nyil more consistent MISC category
    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense of NE: \\ \textit{WorldCom} (\texttt{ORG}) \nyil \textit{WorldCom-botrány} ['WorldCom scandal'] (\texttt{O})
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Data description}

\begin{frame}{Data format}

@@

\begin{frame}{Data description}

\begin{itemize}
\item available under CC-BY-SA 3.0 Unported License
\item freely downloadable from {\tt http://hlt.sztaki.hu/resources/} 
\item also distributed through META-SHARE (from August)
\end{itemize}

@@

\begin{frame}{Silver standard corpora}

@@

\begin{itemize}
\item for less resourced languages they can serve as training corpora in lieu of gold standard datasets
\item they can serve as training sets for domains differing from newswire
\item they can be sources of huge entity lists
\item they can be sources of feature extraction
\end{itemize}

@@

\begin{itemize}
\item gazetteer features
\item sentence start and end position
\item Boolean-valued orthographic properties of the word form
\item string-valued properties of the word form
\item morphological information
\end{itemize}

@@

\begin{frame}{Corpus size and NE density}

@@

quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets

@@

\begin{nix}
our experiments were performed with a sample of 3.5 million tokens
\end{nix}

@@

\begin{frame}{English results}

@@

\begin{frame}{Hungarian results}

@@

\begin{itemize}
\item F-measures for Hungarian are higher 
\begin{itemize}
\item the MISC category is more consistent
\item the Szeged NER corpus is highly accurately tagged 
\end{itemize}
\item the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
\end{itemize}

@@

\section{Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
 \vitem we presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
  \vitem we applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
  \vitem the huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
  \vitem silver standard corpora can improve NER accuracy in more ways
  \vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
  \end{itemize}
  \vfill
\end{frame}

@@

{\LARGE Thank you for your attention!}

@@

emails: {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\

>>>>

\newcommand{\vitem}{\item \vspace{4pt}}
\newcommand{\nyil}{$\rightarrow$\ }
\newcommand{\nagytilde}{$\sim$}
\newtheorem{Examplee}{Example}[section]

@@

@@

The standard datasets are
\smallskip
\begin{itemize}
\item highly domain-specific
\item restricted in size
\end{itemize}

To get a bigger training corpus:
\smallskip
\begin{itemize}
\item merging datasets \nyil problems of combining different tagsets
\item manual annotation \nyil not cost-effective
\end{itemize}

@@

\begin{frame}{Reducing the Annotation Cost}

@@

Reducing the annotation cost:
\smallskip
\begin{itemize}
\item to generate the resource automatically 
\item to use collaboratively constructed resources (e.g.~Wikipedia, DBpedia)
\end{itemize}

@@

\begin{nix}
Combining these approaches: \\ automatically generating NE tagged corpora from Wikipedia
\end{nix}

@@

Improving the accuracy of NER with WP:
\smallskip
\begin{itemize}
\item extracting gazetteers containing person names, locations and organizations 
\item extracting the knowledge embodied in WP and incorporating it in NER learning as features
\item automatic creation of training data from WP
\begin{itemize}
\item by using the category structure of WP
\item by using the category head nouns and the opening sentences of articles
\end{itemize} 
\end{itemize}

@@

\begin{frame}{Our Approach}

\begin{nix}
Mapping the DBpedia ontology classes to standard NE tags, and assign them to WP entities
\end{nix}

@@

Our method is largely language-independent \nyil sufficiently large corpora for less resourced languages can be built

@@

\begin{nix}
This is the first automatically annotated corpus for Hungarian 
\end{nix}

@@

\subsection*{Motivation}
\begin{frame}{Motivation}
  \bigskip

  \begin{itemize}
  \item Our goal is to create a large NE annotated corpus, automatically
        generated from WP articles.
  \smallskip
  \vitem Our method is similar to Nothman et al. (2008)
  \vitem The process consists of four steps:
    \begin{enumerate}
    \vitem Classify WP articles into entity classes.
    \vitem \textit{Parse WP and split articles into sentences.}
    \vitem Label named entities in the text.
    \vitem Select the sentences for inclusion in the corpus.
    \end{enumerate}
  \end{itemize}
\end{frame}

\subsection{Articles to Entities}
\begin{frame}{Articles to Entities}
  \bigskip
  Our method is based on thinking of WP articles as NEs.
  \smallskip
  \begin{itemize}
  \vitem Usual approach: semi-supervised learning (infoboxes, first sentence)
  \vitem Our approach: create entity list from DBpedia
    \begin{itemize}
    \vitem OWL ontology with 320 classes
    \vitem We take the most specific class
    \vitem Manual mapping to CoNLL categories that exploits inheritence
    \end{itemize}
  \item Decreased coverage, increased precision
  \end{itemize}
\end{frame}

\subsection{Named Entity Tagging}
\begin{frame}{Named Entity Tagging}
  \bigskip
  The entity list from DBpedia is used for NE tagging.
  \smallskip
  \begin{itemize}
  \vitem Entity identification:
    \begin{itemize}
    \vitem Look up linked pages in the entity list
    \vitem References to the current page, if in the list
    \end{itemize}
  \vitem NE tagging: from the DBpedia--CoNLL category mapping
  \vitem Exceptions:
    \begin{itemize}
    \vitem Not entities: common nouns, date expressions, etc.
    \vitem Unmarked entities: link to already identified entities
    \vitem Special cases: derived words, first word in a sentence, etc.
    \end{itemize}
  \end{itemize}
\end{frame}

\subsection{Sentence Filtering}
\begin{frame}{Sentence Filtering}
  \bigskip
  We only include fully tagged sentences in the corpus.
  \smallskip
  \begin{itemize}
  \vitem Sentences that contain unknown links / uppercase words
  \vitem To filter incomplete sentences:
    \begin{itemize}
    \vitem Sentences without an EOS punctuation mark
    \vitem Sentences without a predicate (verb)
    \end{itemize}
  \end{itemize}
\end{frame}

@@

%TODO: the "Hungarian Corpus" phrase must appear in the frame title 
\begin{frame}{Differences from English}
  \bigskip
  \begin{itemize}
  \vitem Language processing tools
    \begin{itemize}
    \vitem Inflectional morphology: different set of tools were needed 
    \vitem We developed a flexible framework that enables the plugging of any NLP tool
    % TODO: what kind of linguistic tools are needed: table?
    \end{itemize}
  \vitem Wikipedia size and coverage
    \begin{itemize}
    \vitem Hungarian is medium-size language: \nagytilde 200k articles vs.~\nagytilde 4M English pages
    \vitem Cross-references may point to English pages
    \vitem Templates are in Hungarian: per-language template configuration
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Differences from English -- cont.}
  \bigskip
  \begin{itemize}
  \vitem DBpedia coverage
    \smallskip
    \begin{itemize}
    \vitem Internationalized chapters for big languages only
    \vitem "Hungarian" DBpedia list contains English articles % Nem egyértelmű
      \begin{itemize}
      \vitem Hungarian equivalents must be added to the list % Added: lásd előző oldal 2/2
      \vitem Hungarian entities without an English page are missing:
             the NER language model might not be accurate
      \vitem Future work: extend the list with the missing Hungarian entities
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Differences from English -- cont.}
  \bigskip
  \begin{itemize}
  \vitem NE conventions
    \smallskip
    \begin{itemize}
    \vitem Szeged NER guideline: similar to CoNLL
    \vitem Common nouns in Hungarian: names of languages, religions, months, etc. \nyil more consistent MISC category
    \vitem NEs can be subject of compounding \nyil the joint common noun can modify the original sense and category of the NE:
    \end{itemize}
  \end{itemize}
  \bigskip
  \begin{Examplee}
  \small{WorldCom {\normalfont(\texttt{ORG})} \nyil WorldCom-botrány ['WorldCom scandal'] {\normalfont(\texttt{O})}}
  \end{Examplee}
\end{frame}

\section{Data Description}

\begin{frame}{Data Format}

@@

\begin{frame}{Data Availability}
\bigskip
The corpora are
\begin{itemize}
\vitem available under CC-BY-SA 3.0 Unported License
\vitem freely downloadable from \url{http://hlt.sztaki.hu/resources/} 
\vitem also distributed through META-SHARE (from August)
  \begin{itemize}
  \vitem an open, distributed facility for exchanging and sharing resources
  \vitem one of the lines of action of META-NET, a Network of Excellence funded
        by the European Commission
  \vitem \url{http://www.meta-net.eu/meta-share}
  \end{itemize}
\end{itemize}

@@

\begin{frame}{Silver Standard Corpora}

@@

\begin{itemize}
\vitem for less resourced languages they can serve as training corpora in lieu of gold standard datasets
\vitem they can serve as training sets for domains differing from newswire
\vitem they can be sources of huge entity lists
\vitem they can be sources of feature extraction
\end{itemize}

@@

\begin{itemize}
\vitem gazetteer features
\vitem sentence start and end position
\vitem Boolean-valued orthographic properties of the word form
\vitem string-valued properties of the word form
\vitem morphological information
\end{itemize}

@@

\begin{frame}{Corpus Size and NE Density}

@@

Quite low NE density \nyil sentences without NEs were filtered out from the evaluation datasets

@@

\begin{nix}
Our experiments were performed with a sample of 3.5 million tokens
\end{nix}

@@

\begin{frame}{English Results}

@@

\begin{frame}{Hungarian Results}

@@

\begin{itemize}
\vitem F-measures for Hungarian are higher 
\begin{itemize}
\vitem the MISC category is more consistent
\vitem the Szeged NER corpus is highly accurately tagged 
\end{itemize}
\vitem the Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
\end{itemize}

@@

\section{Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
  \vitem We presented freely available NE tagged corpora for English and Hungarian, fully automatically generated from WP
  \vitem We applied a new approach: mapping DBpedia ontology classes to standard CoNLL NE tags, and assigning them to WP entities
  \vitem The huge amount of WP articles opens the possibility of building large enough corpora for less resourced languages such as Hungarian
  \vitem Silver standard corpora can improve NER accuracy in more ways
  \vitem The Hungarian corpus can serve as a training corpus to build NE taggers for non-newswire domains
  \end{itemize}
  \vfill
\end{frame}

@@

{\huge Thank you for your attention!}

\bigskip
\bigskip
\bigskip

emails: \\ {\tt simon.eszter@nytud.mta.hu \\ nemeskey.david@sztaki.mta.hu} \\

@@
